Comparative Analysis of the wav2vec 2.0 Feature Extractor

### 2023-03-09 main.tex
squeeze space
make blind for review
nitpicks

% **************************************
% *    DOUBLE-BLIND REVIEW SETTINGS    *
% **************************************
% Comment out \interspeechcameraready when submitting the 
% paper for review.
% If your paper is accepted, uncomment this to produce the
%  'camera ready' version to submit for publication.
\interspeechcameraready

@@

\subsection{Acoustic Model}
The \gls{AM} mainly consists of \gls{VGG} \addref blocks for downsampling and 12 \conformer blocks.
Its configuration mostly follows \cite{zeineldeen2022conformer} and \cite{zhou2022efficient}.
The \gls{VGG} downsampling uses three 3x3-convolutional layers with 32, 64 and 64 channels, respectively.
In total, a subsampling factor of 4 in the time dimension is achieved by the strided convolutions so that the \conformer operates on frames with a shift of \SI{40}{\milli\second}.
The feature dimension is reduced by a factor of 2 due to max pooling and then multiplied by the number of channels in the last convolution resulting in a total increase by factor 32 in our case.
After a linear projection to a 512-dimensional representation and dropout of 0.1, the 12 \conformer blocks with an inner dimension of 512, swapped convolution and multi-headed self-attention modules and 8 heads are applied.

@@

The number of parameters for Gammatone and log Mel features refers to the \gls{FIR} and Mel filterbanks, respectively, that we use in our implementation even if they are not trainable.
The differences in the total number of parameters are influenced by the size of the \fe, but also of the linear layer before the \conformer which grows large as the feature dimension increases.
\input{tables/features_general}

\subsection{Dissecting the wav2vec 2.0 Feature Extractor}
\label{sec:w2v_components}
It is striking that the \wvtwo \fe uses two orders of magnitude more parameters than the \gls{SC} architecture.
To understand where \wvtwo uses the them and how much they contribute to the \gls{FE}'s performance, we run it with different configurations.
The results for different widths and depths are shown in \refTab{table:features_w2v_size}.
The kernel sizes and strides were chosen as (10, 3, 3, 3, 3, 2), (5, 2, 2, 2, 2, 2) for 6 layers, (10, 6, 3, 3, 3), (5, 4, 2, 2, 2) for 5 layers, (10, 6, 6, 3), (5, 4, 4, 2) for 4 layers, (20, 6, 6), (10, 4, 4) for 3 layers and (32, 20), (16, 10) for 2 layers.
The results show that decreasing the inner dimension deteriorates the performance, however, a larger dimension does not improve over the 512-dim baseline.
In contrast, no major effect can be observed regarding the number of layers.
We also tried a variant with only one layer using kernel size 320 and stride 160 which did not converge.
\input{tables/features_w2v_size}
Removing the final projection with about 394k parameters has no significant impact.
%The final projection, which comprises about 394k parameters, additionally improves the performance slightly on \devother while degrading minimally on the \textit{test} sets as shown in \refTab{table:features_w2v_proj}.
%\input{tables/features_w2v_proj}

@@

\subsection{Frequency Response of Learned Filters}
\label{sec:freq_resp}
To analyze the learned filters of the neural \fes, we first plot the frequency response of the filters in the first layer which operates on the raw waveform.
As the order of learned filters in a convolutional layer is arbitrary, we sort the filters by the peak value of the frequency response and by the upper and lower \SI{3}{\decibel} cutoff frequencies as second and third criteria, respectively.
The plots are depicted in \refFig{fig:first_layer}\,(a)-(d), which share the same color range.

>>>>

% **************************************
% *    DOUBLE-BLIND REVIEW SETTINGS    *
% **************************************
% Comment out \interspeechcameraready when submitting the 
% paper for review.
% If your paper is accepted, uncomment this to produce the
%  'camera ready' version to submit for publication.
% \interspeechcameraready

@@

\subsection{Acoustic Model}
The \gls{AM} mainly consists of 3 VGG blocks for downsampling and 12 \conformer blocks.
Its configuration mostly follows \cite{zeineldeen2022conformer} and \cite{zhou2022efficient}.
The VGG downsampling uses three 3x3-convolutional layers with 32, 64 and 64 channels, respectively.
In total, a subsampling factor of 4 in the time dimension is achieved by the strided convolutions so that the \conformer operates on frames with a shift of \SI{40}{\milli\second}.
The feature dimension is reduced by a factor of 2 due to max pooling and then multiplied by the number of channels in the last convolution resulting in a total increase by factor 32 in our case.
After a linear projection to a 512-dimensional representation and dropout of 0.1, the 12 \conformer blocks with an inner dimension of 512, swapped convolution and multi-headed self-attention modules and 8 heads are applied.

@@

The number of parameters for Gammatone and log Mel features refers to the \gls{FIR} and Mel filterbanks, respectively, that we use in our implementation even if they are not trainable.
The differences in the total number of parameters are influenced by the size of the \fe, but also of the linear layer before the \conformer which grows large as the feature dimension increases.
\input{tables/features_general}
\vspace{-0.3cm}

\subsection{Dissecting the wav2vec 2.0 Feature Extractor}
\label{sec:w2v_components}
It is striking that the \wvtwo \fe uses two orders of magnitude more parameters than the \gls{SC} architecture.
To understand where \wvtwo uses the them and how much they contribute to the \gls{FE}'s performance, we run it with different configurations.
The results for different widths and depths are shown in \refTab{table:features_w2v_size}.
The kernel sizes and strides were chosen as (10, 3, 3, 3, 3, 2), (5, 2, 2, 2, 2, 2) for 6 layers, (10, 6, 3, 3, 3), (5, 4, 2, 2, 2) for 5 layers, (10, 6, 6, 3), (5, 4, 4, 2) for 4 layers, (20, 6, 6), (10, 4, 4) for 3 layers and (32, 20), (16, 10) for 2 layers.
The results show that decreasing the inner dimension deteriorates the performance, however, a larger dimension does not improve over the 512-dim baseline.
In contrast, no major effect can be observed regarding the number of layers.
We also tried a variant with only one layer using kernel size 320 and stride 160 which did not converge.
Removing the final projection with about 394k parameters has no significant impact.
%The final projection, which comprises about 394k parameters, additionally improves the performance slightly on \devother while degrading minimally on the \textit{test} sets as shown in \refTab{table:features_w2v_proj}.
\input{tables/features_w2v_size}
%\input{tables/features_w2v_proj}

@@

\subsection{Frequency Response of Learned Filters}
\label{sec:freq_resp}
To analyze the learned filters of the neural \fes, we first plot the frequency response of the filters in the first layer which operates on the raw waveform.
As the order of learned filters in a convolutional layer is arbitrary, we sort the filters by the peak value of the frequency response and by the upper and lower \SI{3}{\decibel} cutoff frequencies as second and third criteria.
The plots are depicted in \refFig{fig:first_layer}\,(a)-(d), which share the same color range.
